Multiple regression: TRELIS-Q WP5 data

### 2024-08-20 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
Added ICC and updated graph

---
title: "Multiple regression: TRELIS-Q WP5 data"
author: "Niklas Karlsen"
date: "2024-04-02"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

@@

@@

# cluster-variable
# kilde: https://www.r-bloggers.com/2021/05/clustered-standard-errors-with-r/
library(estimatr)                           # for cluster corrected SE
#library(sandwich)
#library(lmtest)
```

@@

```{r}
ggplot(dc, aes(x=TPACK, y=var, fill=studiested2)) + 
  geom_bar(stat="identity", position="dodge") +
  #scale_fill_manual(values=viridis(3)) + 
  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
                  linewidth=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) +
  ggtitle("PCK, TK og TPACK fordelt på studiested") +
  guides(fill="none") +
  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
#  coord_cartesian(ylim = c(1,5)) +
  ylab("verdi") +
  xlab("") +
  theme_bw() +
  facet_wrap(~studiested2)
```

@@

@@

>>>>

---
title: "Multiple regression: TRELIS-Q WP5 data"
author: "Niklas Karlsen"
date: "2024-08-13"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

@@

TODO

* endre navn til paper 2

Changelog:

* 13.8.2024: added check of ICC
* 16.5.2024: added plot for NFSUN2024

@@

# cluster-variable
# kilde: https://www.r-bloggers.com/2021/05/clustered-standard-errors-with-r/
library(estimatr)                           # for cluster corrected SE
#library(sandwich)
#library(lmtest)

library(nlme)                               # for ICC - variance HLM
```

@@

For NFSUN2024:

```{r}
# anonymise institutes
data_clean$studiested3 <- factor(data_clean$studiested, 
                                 level=c("HVL","UiT","Nord universitet","Høgskolen i Innlandet","OsloMet"),
                                 label=c("I1", "I2", "I3", "I4", "I5"))

dc1 <- summarySE(data_clean, measurevar="TK", groupvars="studiested3")
dc2 <- summarySE(data_clean, measurevar="PCK", groupvars="studiested3")
dc3 <- summarySE(data_clean, measurevar="TPACK", groupvars="studiested3")

# add column med navn
dc["TPACK"] <- c("TK","TK","TK","TK","TK",
                 "PCK","PCK","PCK","PCK","PCK",
                 "TPACK","TPACK","TPACK","TPACK","TPACK")
```

```{r}
stolpediagram <- ggplot(dc, aes(x=studiested3, y=var, fill=studiested3)) + 
  geom_bar(stat="identity", position="dodge") +
  #scale_fill_manual(values=viridis(3)) + 
  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
                  linewidth=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) +
  ggtitle("Mean and standard deviation for PCK, TK and TPACK") +
  guides(fill="none") +
  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
#  coord_cartesian(ylim = c(1,5)) +
  ylab("Value") +
  xlab("Institute") +
  theme_bw() +
  facet_wrap(~TPACK)

```{r}
# save plot
if (plot_to_file) {
  ggsave("../output/NFSUN2024_stolpediagram.png",dpi=300) #height=8,width=15)
  }

```{r eval=F}
ggplot(dc, aes(x=TPACK, y=var, fill=studiested2)) + 
  geom_bar(stat="identity", position="dodge") +
  #scale_fill_manual(values=viridis(3)) + 
  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
                  linewidth=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) +
  ggtitle("PCK, TK og TPACK fordelt på studiested") +
  guides(fill="none") +
  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
#  coord_cartesian(ylim = c(1,5)) +
  ylab("verdi") +
  xlab("") +
  theme_bw() +
  facet_wrap(~studiested2)
```

@@

### ICC

ICC: Christophersen (2018, p. 109-111)

```{r ICC}
# variance decomposition
null.mod <- lme(TPACK ~ 1, random =~ 1|studiested, data = data_clean)
vcn <- VarCorr(null.mod)
vcn
```

The variance of indviduals is given by the residual. The variance of institutions is givene by the intercept. If the variance of the intercept is larger than 0.05 of the total variance (intercept + residual), then the institutional variance contributes to explain the outcome.

### Normality

@@

### Linearity
